---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# hydrogeologicalghana

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)

<!-- badges: end -->

The goal of hydrogeologicalghana is to provide data on boreholes collected to study the hydrogeological dynamics in the greater Accra and Ashanti regions of Ghana.

## Installation

You can install the development version of hydrogeologicalghana from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/hydrogeologicalghana")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "gt", "kableExtra"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

1.  Click Download CSV. A window opens that displays the CSV in
    your browser.
2.  Right-click anywhere inside the window and select "Save Page As...".
3.  Save the file in a folder of your choice.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/openwashdata/hydrogeologicalghana/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data

The package provides access to measurements of pumping tests that were systematically conducted at boreholes to assess crucial hydrogeological parameters, specifically focusing on pumping rates and drawdown.

```{r}
library(hydrogeologicalghana)
```

### hydrogeologicalghana

The dataset `hydrogeologicalghana` contains data about pumping tests that were systematically conducted at boreholes to assess crucial hydrogeological parameters, specifically focusing on pumping rates and drawdown in the greater Accra and Ashanti regions of Ghana. It has
`r nrow(hydrogeologicalghana)` observations and `r ncol(hydrogeologicalghana)`
variables

```{r}
hydrogeologicalghana |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "hydrogeologicalghana.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Example

```{r}
library(hydrogeologicalghana)
library(dplyr)

# Summary statistics by region
region_summary <- hydrogeologicalghana |>
  group_by(region) |> 
  summarise(
    count = n(),
    avg_depth = mean(depth, na.rm = TRUE),
    max_pump_rate = max(pumping_rate, na.rm = TRUE),
    min_water_level = min(static_water_level, na.rm = TRUE)
  )

# Print the formatted table
kable(region_summary, 
      caption = "Summary Statistics by Region", 
      align = "c",
      col.names = c("Region",
                    "Count",
                    "Average depth",
                    "Max. pump rate",
                    "Min. water level"))
```

```{r}
library(hydrogeologicalghana)
library(ggplot2)

# Scatter plot
scatter_plot <- hydrogeologicalghana |>
  ggplot(aes(x = depth, y = static_water_level, color = region)) +
  geom_point() +
  labs(title = "Scatter plot of depth vs. water level by region",
       x = "Depth (m)",
       y = "Water Level (m)") +
  theme_minimal()

# Display the scatter plot
print(scatter_plot)
```

## License

Data are available as
[CC-BY](https://github.com/openwashdata/%7B%7B%7Bpackagename%7D%7D%7D/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("hydrogeologicalghana")
```
